## Initial Item Analysis

```{r, include=FALSE}
knitr::opts_chunk$set(comment = "", echo=FALSE, warning = FALSE, message = FALSE, tidy.opts=list(width.cutoff=60), tidy = TRUE)
knitr::opts_knit$set(verbose = FALSE)
```

In Table X we report classical statistics for the 66 SHAS items. We sorted the items by their mean. We flagged six items due to extreme values of skew (less than -2 or greater than +2) or kurtosis (less than -7 or greater than +7) and item-total correlations below 0.50. These included items EQ17, EQ27, EQ41, EQ46, EQ55, and EQ60. Category response proportions show that most people responded "Never" to these items. Thus these items do not apply to the vast majority of people in this population. We removed them from the item pool.

```{r, echo=FALSE}

# RUN CTT STATS, BUILD ITEM TABLE
rm(list=ls())

# LOAD DATA
all <- read.csv("data/all.csv")
all <- all[,-1]
source("code/ApplyAllLabels.R")
library(psych)
library(ltm)

# DESCRIPTIVES
descriptives <- as.data.frame(psych::describe(all[,1:66]))
descriptives <- subset(descriptives, select = -c(vars, trimmed, se, median, mad, min, max, range))

# CATEGORY FREQUENCIES
catfreqs <- ltm::descript(all[,1:66])
catfreqs <- as.data.frame(catfreqs[["perc"]])
#names(catfreqs) <- c('1' = "Never", '2' = "Once or twice", '3' = "Sometimes", '4' = "Often", '5' = "All the time")

# ITEM TOTAL CORRELATIONS
a <- psych::alpha(all[,1:66])
item.totals <- as.data.frame(a[["item.stats"]][["raw.r"]])
names(item.totals) <- c('a[["item.stats"]][["raw.r"]]'= "Item-total correlation")

# BUILD TABLE
item.table <- cbind(catfreqs, descriptives, item.totals)
item.table$flag <- with(item.table, ifelse(item.table$skew > 2 | item.table$skew < -2 | item.table$kurtosis < -7 | item.table$kurtosis > 7 | item.table$'Item-total correlation' < .5, "*", ""))
item.table2 <- item.table[order(item.table$mean),]

# CLEAN UP
rm(all,descriptives,catfreqs,a,item.totals,item.table)
```

```{r, echo=FALSE}

# RENDER CTT ITEM STATS TABLE
library(kableExtra)
item.table2 %>%
   kbl(digits=2, align = "ccccccccc", longtable = T, booktabs = T) %>%
   kable_styling(latex_options = c("repeat_header")) %>%
   add_header_above(c(" ", "Category response proportion" = 5, " " = 7)) %>%
   column_spec(1, width="2cm")
```

```{r}
# SAVE NEW ITEM FILE
all <- read.csv("data/all.csv")
items2 <- all[ ,1:67]
items2 <- items2[,-1]
items2 <- items2[,-c(5,15,28,33,41,46)]
write.csv(items2,"data/items2.csv")
rm(list=ls())
```
