
```{r, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(comment = "", warning = FALSE, echo = FALSE,  message = FALSE, tidy.opts=list(width.cutoff=60), tidy = TRUE)
knitr::opts_knit$set(verbose = FALSE)
```

#### Local Independence

To check the assumption of local independence, we used the Q3 statistic (Yen, 1984). The Q3 statistic index criteria are that the raw residual correlation between pairs of items should never exceed 0.10 (Marais & Andrich, 2008).

```{r q3, echo=FALSE}
# Q3 STATISTIC
library(sirt)
items <- read.csv("data/ditems_l.csv") # grab the data
mod <- sirt::rasch.mml2(items, progress = FALSE) # estimate the Rasch model
mod.wle <- sirt::wle.rasch(dat=items, b=mod$item$b) # estimate WLEs
mod.q3 <- sirt::Q3(dat=items, theta=mod.wle$theta, b=mod$item$b ) # estimate Q3
```

```{r q3-plot, fig.cap='Correlations matrix of Item Residuals'}
# plot Q3 statistics
I <- ncol(items)
image( 1:I, 1:I, mod.q3$q3.matrix, col=gray( 1 - (0:32)/32),
       xlab="Item", ylab="Item")
abline(v=c(53,66)) # borders for testlets
abline(h=c(53,66))
```

Figure \@ref(fig:q3-plot) is a plot of the matrix of correlations between item residuals. Each square depicts a correlation. The squares are shaded in grey with the lightest shade indicating the weakest correlations. As hoped, the vast majority of squares are very light, indicating weak correlations among the item residuals.

```{r}
rm(list = ls())
```
