
```{r, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(comment = "", warning = FALSE, echo = FALSE,  message = FALSE, tidy.opts=list(width.cutoff=60), tidy = TRUE)
knitr::opts_knit$set(verbose = FALSE)
```

## Graded Response Model

The fourth and final model we consider is the Graded Response Model (). The difference here allows the discrimination parameter to vary. 

```{r grm-run}
# ESTIMATE GRM MODEL
library(knitr)
library(kableExtra)
library(stats4)
library(lattice)
library(mirt)
items <- read.csv("data/items_l.csv")
source("code/ApplyItemLabels.R")
model.grm <- 'GRADED RESPONSE MODEL = 1-66'
results.grm <- mirt(data=items, model=model.grm, itemtype="graded", SE=TRUE,  technical = list(NCYCLES = 10000), verbose=FALSE) # estimate model
```

```{r grm-item-stats}
# GET ITEM PARAMETERS
coef.grm <- coef(results.grm, IRTpars=TRUE, simplify=TRUE)
item.pars.grm <- as.data.frame(coef.grm$items)

# ITEM FIT
item.fits.grm <- as.data.frame(mirt::itemfit(results.grm, 'infit'))
item.fits.grm <- item.fits.grm[,-1]
items.tbl.grm <- cbind(item.pars.grm,item.fits.grm)

# i <- as.data.frame(itemfit(results.grm, 'infit'))
# i2 <- i[order(i$outfit),]
# misfits <- subset(i2, outfit > 1.5 | outfit < 0.5 | infit > 1.5 | infit < 0.5)
# i <- items.grm[order(items.grm$a),]
# items.pcm$deltab1b2 <- items.pcm$b2 - items.pcm$b1 # look for reversals b1 to b2
# items.pcm$deltab2b3 <- items.pcm$b3 - items.pcm$b2 # look for reversals b2 to b3
# items.pcm$deltab3b4 <- items.pcm$b4 - items.pcm$b3 # look for reversals b3 to b4
# items.pcm$avg <- rowMeans(items.pcm[ ,6:8])
# rv <- subset(items.pcm, deltab1b2 < 0 | deltab2b3 < 0 | deltab3b4 <0) # subset items with reversals
# rv <- rv[order(rv$avg),]
# rv2 <- subset(rv, select = -c(deltab1b2:deltab3b4))

```

```{r grm-items-tbl}
library(kableExtra)
items.tbl.grm %>%
  kable(digits = 3, align = "ccccccccc", caption = 'Graded Response Model Item Paramters and Fit Statistics') %>%
  kable_styling(bootstrap_options = "condensed", font_size = 12, fixed_thead = T)
```

Again, plots of the option characteristic curves help convey this information visually. The following plots illustrate three items with very different sets of curves. The first item illustrates the intended pattern of ordered thresholds. The second and third items show pronounced reversals. 

```{r grm-a, fig.cap='Option Characteristic Curves for Graded Response Model'}
# PCM ITEM PLOTS
library(directlabels)
a <- plot(results.grm, type = 'trace', which.items = c(41,37,42),
     main = "", par.settings = simpleTheme(lty=1:4,lwd=2),
     auto.key=list(points=FALSE,lines=TRUE, columns=4))
direct.label(a, 'top.points')
```

The items fit.

The obtained RMSEA value = .079 (95% CI[.078, .080]) and SRMSR value = .051 suggest that data do not fit the model reasonably very well using suggested cutoff values of RMSEA <= .06 and SRMSR <= .08 as suggested guidelines for assessing fit. The CFI = .97 exceeded the recommended .95 threshold.

```{r grm-model-fit}
# MODEL FIT
grm.model.fit <- as.data.frame(M2(results.grm, type = "C2", calcNULL = FALSE))
row.names(grm.model.fit)[row.names(grm.model.fit) == "stats"] <- "GRM"
grm.model.fit %>%
  kable(digits = 3, align = "ccccc", caption = 'Graded Response Model Fit Statistics') %>%
  kable_styling(bootstrap_options = "condensed", font_size = 12, fixed_thead = T)
```

```{r}
save(results.grm, file = "data/results.grm.RData")
write.csv(grm.model.fit, "data/grm.model.fit.csv")
rm(list = ls())
```
